Supplementary data for: \\textit{`MSnbase`, efficient and elegant R-based processing and visualisation of raw mass spectrometry data}

### 2020-06-26 large_experiment_processing.Rmd
Add section with general usage

>>>>

## Basic MS data access functionality

Before evaluating the `MSnbase` performance on the large data set we provide
first some general description of the `MSnbase` data classes in `MSnbase` basic
data access operations. MS data from raw data files in mzML, mzXML, mzData or
netCDF format is represented by the `MSnExp` object which organizes the spectra
from the original files in a one-dimensional list. Functions like `rtime` and
`msLevel` allow to extract the retention time and MS level, respectively, for
each spectrum and return a `numeric` (or `integer`) vector with the same length
than the number of spectra in the `MSnExp`. In the example below we use the
`rtime` function to extract the retention times for each spectrum.

```{r rtime}
rts <- rtime(ms_dsk)
length(rts)
head(rts)
```

The `fromFile` function can be used to determine from which file (sample) a
specific spectrum in the `MSnExp` object is. This function returns an `integer`
vector, same length than spectra in the experiment, with the file index; the
file names can be accessed with the `fileNames` method. A `MSnExp` object can be
subsetted with `[` and e.g. the index of the spectra that should be retained. In
the code block below we subset our `ms_dsk` object to contain only spectra from
the 3rd file.

```{r subset}
one_file <- ms_dsk[fromFile(ms_dsk) == 3]
length(one_file)
```

Note however that there are also dedicated *filter* functions to subset an
`MSnExp` object such as `filterFile`, `filterMsLevel`, `filterRt`, `filterMz`,
`filterPrecursorMz` or `filterIsolationWindow`. In the example below we use the
`filterRt` function to further subset our data to contain only spectra acquired
within a certain time range.

```{r filterrt}
one_file <- filterRt(one_file, rt = c(40, 300))
length(one_file)
```

As mentioned above, the `MSnExp` object is comparable with a list of
spectra. Thus, to extract a single spectrum from it we can use `[[`. This will
return an object of type `Spectrum` which encapsules/represents all information
of the measured spectrum (i.e. m/z and intensity values as well as metadata
information). In the example below we extract the 15th spectrum from our data
subset and access its m/z values with the `mz` function.

```{r}
sp <- one_file[[15]]
mz(sp)
```

This particular spectrum has thus only 3 peaks.

Note that m/z or intensity values can also be directly extracted from the
`MSnExp` object as shown in the example below. The result will be a `list` of
`numeric` vectors, each element representing the m/z values for each spectrum in
the object.

```{r}
mzs <- mz(one_file)
class(mzs)
length(mzs)
```

In addition, it is also possible to extract all m/z and intensity values from an
`MSnExp` object as a `data.frame` as shown in the code block below, but this is
not suggested, since it loads all the data into memory and all MS spectrum
metadata such as MS level or precursor m/z get lost.

```{r data.frame}
df <- as(one_file, "data.frame")
head(df)
nrow(df)
```

Note that for all these operations it is irrelevant whether an in-memory or
on-disk backend was used. In general it is advisable to use the on-disk backend
especially for experiments with more than ~ 50 files.
